Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-04 Manuscript/manuscript.tex
Corrected section Stoichiometry.

\subsection{Structure and affinity of cation binding to POPC membrane}

@@

\subsection{Binding stoichiometry to POPC membrane}

@@

Simple binding models have been previous used to interpreted 
the same experimental data, which is used in this work to validate
the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}~\cite{altenbach84,macdonald87}.
The experimental results from the lipid headgroup order parameters and from the
atomic absorption spectroscopy were best
explained by the ternary complex binding model with a binding
stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
However, also a Langmuir adsorption model (i.e. \ce{Ca^{2+}}:POPC stoichiometry of 1:1) 
was successfully fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
The results based on these binding models and the measured headgroup order parameter changes
are also inluded in the tabulated experimental binding constants~\cite{marsh13}.

Since the same experimental data is reproduced by the ECC-lipids model,
it can be used to give a more versatile interepration of the binding stoichiometry than
the previously used simple binding models~\cite{altenbach84,macdonald87}. To directly evaluate the
stoichiometry from simulations we calculated the relative propensities of \ce{Ca^{2+}}:POPC
complexes with various stoichiometries by counting the individual lipid
molecules within the distance of $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion.
The results from POPC simulation with 285~mM concetration of CaCl$_2$ in
Fig.~\ref{fig:cacl_complexes} show the largest propensity (41\%) for the
ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
the probabilities for the complexes with the stoichimetries of 1:1~(25\%)~and
1:3~(34\%)~are only slightly lower,
suggesting more complex binding stoichiometry than the simple ternary complex model.
%The observed propensities of
%$41\%$ for two lipids,
%$25\%$ for one lipid, and
%$34\%$ for three lipids 
However, one calcium binds roughly to two lipids on average also in simulations,
because the probabilities of forming the complexes with 1 or 3 lipids are almost equal.  
This probably explains the successfull fit of the ternary complex model to the
experimental data. 
Complexes with more than three lipids per one calcium ion were not observed in 
simulations with the ECC-lipids model.
\todo{I think that we either have to do the analysis in SI properly,
  i.e., using the real $C_I$ from simulations or remove the sentence below and
the related concent from SI.}
Ternary complex model also provides a good fit 
to our simulations with ECC-lipids 
(see Fig.~\ref{fig:cacl-bind} in SI and its caption for details). 
\begin{figure}[tb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
  \caption{\label{fig:cacl_complexes}
      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
      with a certain number of POPC lipids. 
      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
  }
\end{figure}

Also the probabilities of different complexes formed by \ce{Na+} ions and POPC
analyzed from the ECC-lipids model simulation with 1 M concentration of NaCl are
shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
probability is largest (67\%) for 1:1 complex, significantly smaller (29\%)
for 1:2 complexes and very small (4\%) for 1:3 of \ce{Na+}:POPC complexes.

>>>>

\subsection{Structure and affinity of cation binding to POPC membrane}
\label{sec:affinity}

@@

\subsection{Binding stoichiometry of \ce{Na+} and \ce{Ca^{2+}} cations to POPC membrane}

@@

Simple binding models have been previously used to interpret
the same experimental data \cite{altenbach84,macdonald87} 
which are used in this work to validate the simulation models in Fig. \ref{fig:delta_ordPar_CaCl}.
The experimental results from the lipid headgroup order parameters and from the
atomic absorption spectroscopy were best
explained by a ternary complex binding model with a binding
stoichiometry of one \ce{Ca^{2+}} per two POPC lipids~\cite{altenbach84}.
However, also a Langmuir adsorption model (i.e. \ce{Ca^{2+}}:POPC stoichiometry of 1:1) 
provided a good fit to the experimental data when only low concentrations of \ce{CaCl2} were considered~\cite{macdonald87}.
The results based on these binding models and the measured headgroup order parameter changes
are also inluded in the tabulated experimental binding constants~\cite{marsh13}.

Since the same experimental data is reproduced by the ECC-lipids model,
it can be used to give a more versatile interepration of the binding stoichiometry than
the previously used simple binding models~\cite{altenbach84,macdonald87}. To directly evaluate the
stoichiometry from simulations we calculated the relative propensities of \ce{Ca^{2+}}:POPC
complexes with various stoichiometries by counting the individual lipid
molecules within the distance $0.3\,\mathrm{nm}$ from each \ce{Ca^{2+}} ion
as already used in previous section~\ref{sec:affinity}.
The results from POPC simulation with 285~mM concetration of CaCl$_2$ in
Fig.~\ref{fig:cacl_complexes} show the largest propensity (41\%) for the
ternary complex with the \ce{Ca^{2+}}:POPC stoichiometry of 1:2, but 
the probabilities of complexes with the stoichimetries of 1:1~(25\%)~and
1:3~(34\%)~are only slightly lower
suggesting a more complex binding model than the simple ternary complex model.
%The observed propensities of
%$41\%$ for two lipids,
%$25\%$ for one lipid, and
%$34\%$ for three lipids 
However, the simulation data can also be viewed from such an angle that
one calcium binds roughly to two lipids on average,
because the probabilities of the complexes with 1 or 3 lipids are almost equal.  
This probably explains the success of the ternary complex model fitting well the
experimental data. 
Complexes with more than three lipids per one calcium ion were not observed in 
simulations with the ECC-lipids model.
\todo{Samuli: I think that we either have to do the analysis in SI properly,
  i.e., using the real $C_I$ from simulations or remove the sentence below and
the related concent from SI.\\
Joe: I think it shows that ternary complex makes a good fit to our simulation data. 
The fact that we use a different (and more accurate) estimate of $C_I$ and other quantities does not hamper this statement.}
Ternary complex model also provides a good fit 
to our simulations with ECC-lipids 
(see Fig.~S\ref{fig:cacl-bind} in SI and its caption for details). 
\begin{figure}[tb!]
  \centering
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/stoichiometry_NaCl-CaCl2_comparison_Ecc-lipids.pdf} \\
  \caption{\label{fig:cacl_complexes}
      Relative probabilities of existence of \ce{Na^{+}} or \ce{Ca^{2+}} complexes
      with a certain number of POPC lipids. 
      \ce{Na^{+}} complexes were evaluated from the simulation with 1~M concentration;
      and \ce{Ca^{2+}} complexes were evaluated from the simulation with 287~mM concentration.
  }
\end{figure}

The probabilities of different complexes formed by \ce{Na+} ions and POPC
analyzed from the simulation with the ECC-lipids model at $1\,$M concentration of NaCl are also 
shown in Fig.~\ref{fig:cacl_complexes}. In contrast to calcium, the
probability is largest (67\%) for 1:1 complex, significantly smaller (29\%)
for 1:2 complexes and very small (4\%) for 1:3 of \ce{Na+}:POPC complexes.
